# Holds chunks to be inserted later

---
class: left

### A bad function

  - One example in base R is the **NROW** or **NCOL** functions. They return a value even if the
    input is a vector or single value instead of a data frame:
    
  ```{r nrow-example, eval = TRUE, echo = TRUE}
    df <- data.frame(a = 1:3, b = letters[1:3])
    NROW(df)

    vec <- letters[1:10]
    NROW(vec)
    
    single <- "A"
    NROW(single)
  ```
  - **NEVER** use these functions or be tempted to create such functions.


---
class: center

### R is hard to learn (2)
.left[
Matrices and data.frames are not the same, they are stored internally in two completely different ways
- data.frames are stored as a list of columns (which are vectors)
- matrices are stored as a sequential 1-dimensional vector with dimension information so R knows how to break the vector up into 2-dimensions
- because of this, matrices are much faster for math operations
- data.frames use `names()` to get or set the column names
- matrices use `colnames()` to get or set the column names
  
A matrix is a linear vector
```{r matrix-impl, echo = TRUE, eval = TRUE}
mat <- matrix(1:9, nrow = 3)
print(mat)

lobstr::sxp(mat)
```
]

---
class: center

### R is hard to learn (3)
.left[
A data.frame is a list of column data
```{r dataframe-impl,echo = TRUE, eval = TRUE}
df <- data.frame(x = c(1, 2, 3), y = c(4, 5, 6), z = c(7, 8, 9))
print(df)

lobstr::sxp(df)
```
Some functions designed to work with data.frames will fail with matrices, and vice-versa. Worse, some may succeed but give you the wrong output.

]

---
class: center

### There are no operators

.left[
There are no operators in R, only functions. For example
```{r operators, echo = TRUE, eval = TRUE}
5 + 8 * 10
```
is actually called like this by R
```{r non-operators, echo = TRUE, eval = TRUE}
`+`(5, `*`(8, 10))
```
Here, `` `r '\x60+\x60()'` `` and `` `r '\x60*\x60()'` `` are functions taking two arguments each. All operators are actually functions like this.
To prove these two lines of code are equivalent, look at the Abstract Syntax Trees (ASTs) that R creates when it parses an expression like this:
]
.left[
.pull-left[
```{r ast-operators, echo = TRUE, eval = TRUE}
lobstr::ast({5 + 8 * 10})
```
]
.pull-right[
```{r ast-non-operators, echo = TRUE, eval = TRUE}
lobstr::ast({`+`(5, `*`(8, 10))})
```
]
]

---
class: center

### `if`, `else`, and all left braces (`[`, `(`, and `{`) are functions

.left[
Almost everything in R is a function. This can help you when trying to understand code

```{r simple-if-statement, echo = TRUE, eval = TRUE}
t_or_f <- sample(c(TRUE, FALSE), size = 1)
if(t_or_f){
  x <- -9
  y <- -99
  z <- -999
}else{
  x <- 10
  y <- 100
  z <- 1000
}

paste0("t_or_f = ", t_or_f, ", x = ", x, ", y = ", y, ", z = ", z)
```
Can be written like this:
```{r exotic-if-statement, echo = TRUE, eval = TRUE}
if(t_or_f) `{`(x <- -9, y <- -99, z <- -999) else(`{`(x <- 10, y <- 100, z <- 1000))

paste0("t_or_f = ", t_or_f, ", x = ", x, ", y = ", y, ", z = ", z)
```
This is because `if`, `else`, and `{` are all functions so an `if..else` statement is implemented as this:
`if()`  `` `r '\x60{\x60()'` `` `` `r 'else(\x60{\x60())'` `` and `` `r '\x60{\x60()'` `` has each line of code inside as an argument.
]

---
class: left

### Variable naming syntax

- Inconsistent function names:
  - `names()` vs `colnames()`
  - `row.names()` vs `rownames()`
  - `Sys.time()` vs `system.time()`
  


